Evaluating a simplified forecast model in comparison to the ECDC forecasting hub ensemble

### 2022-08-30 paper/paper.Rmd
grammar check

Real-time evaluation has shown that each of these classes of models may perform well though in general semi-mechanistic models, and models that include human insights have outperformed other approaches during certain periods of time [@REF].

@@

The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within week, reporting periodicity [@REF].

@@

Based on these insights from observing the performance of our own forecast submissions and from the hub ensembles over time here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@REF] with aim of producing ensemble like performance without sacrificing interpretability.
The first simplification we make is to model only weekly data, rather than using daily data and then aggregating, this means that the impact of daily reporting artefacts is mitigated.
It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model this leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.
The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
This reduces computational requirements and removes the need for external information on the delay from infection to report which can potentially be misspecified leading to potentially reduced forecast performance.
However, this sacrifices some of the interpretability of the forecast model as any epidemiological summary statistics we now calculate will be based on reported cases and not latent infections.
As discussed in [@gostic] this leads to varying amounts of bias depending on the epidemic phase. 
The final simplification we make is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@EpiNow2].
This represents a parsimonious approach in that we encode our expectation that the growth rate should vary over time and allow this to influence the forecast but we include only a single lag term, reducing the computational overhead of the model.
In order to model potential unobserved interventions and more general changes in transmission, we include an additional growth rate modifier restricted to be between 0 and 1 that differs depending on if the growth rate is positive or negative (due to potential differing responses when cases are growing or increasing) and that acts in a multiplicative fashion (meaning that larger absolute growth rates are reduced to zero growth more rapidly).
This reflects a simplified interpretation of how the ensemble appears to react to potential future changes by assuming a gradual return to stable incidence. 
The only observation for which we do not make an adaption is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion. 
This choice is motivated by our belief that the underlying transmission process is an exponential one and therefore a count error model (where variance is linked to the mean)  is a sensible choice though this is an area of current research.
We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations.

@@

  The ECDC forecast hub requests forecasts for 1 to 4-week forecast horizon and requires forecasts to use a pre-specified format with 23 quantiles of the predictive probability distribution.
  No restrictions were placed on who could submit forecasts and the hub team actively invited participation from research groups known to be involved with COVID-19 forecasting projects.
  Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent-based models, semi-mechanistic models, expert elicitation models, and ensembles of multiple approaches [@https://covid19forecasthub.eu/community.html].
  Teams submitted forecasts at the latest two days after the complete dataset for the forecast week became available and were allowed to use all data available at the time of submission (i.e including up to two days of data for the current week). 
  The ensemble forecast was constructed by taking the median of all predictive quantiles without the exclusion of any valid submitted forecast (where validity was defined as passing minimal formatting checks by the hub team). 
An ensemble was only produced for locations with at least 3 independent forecast models including the hub baseline model.
  Submitted forecasts and target observations are available from the ECDC forecast hub GitHub repository [@REF]. We provide code in the repository of this study to streamline access.

@@

# Models coverage of forecast dats
forecast_dates_by_models <- metadata |>
  summarise_forecasts_by(var = "target_end_date", by = "model") |>
  DT(order(n)) |>
  DT(, n_per := n / forecast_dates) |>
  DT(, type := fcase(
      model %in% global_models, "global",
      model %in% local_models, "local",
      model %in% partial_models[, model], "partial",
      default = "other"
    )
  )
```

@@

We model the expectation ($\lambda_t$) of reported cases ($C_t$) as an order 1 autoregressive (AR(P)) process by epidemiological week ($t$).
The model is initialised by assuming that the initial reported cases are representative with a small amount of error (2.5%).
We assume a negative binomial observation model with overdispersion $\phi$ for reported cases ($C_t$).

@@

In order to standardise forecasts across forecast locations we first normalised both weekly notified test positive cases and forecast test positive cases by the population in the forecast region scaled to be an incidence rate per 10,000 people.
We then evaluated a subset of forecasts from the following countries Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom visually by forecast horizon (1-4 weeks) for both the ensemble and surrogate model. 
These countries were selected in order to include forecasts based on different numbers and types of submitted forecast models, to be at least partially representative of the full sample of forecast locations, and to include nations for which the authors had a good understanding of local data and transmission dynamics in the study period.

We evaluate forecasts for all locations and horizons quantitatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS).
Both the WIS and CRPS are generalisations of the absolute error in order to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the ECDC forecast hub  [@REF].
We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks).

In order to understand the relative performance of the surrogate model compared to the ensemble model, we calculate the relative performance (rWIS) by dividing the WIS for the surrogate model by the WIS of the ensemble model for all locations and forecast horizons.
In order to maintain the propriety of this score, we do this after first taking the means of scores for the relevant stratification.
We explore relative performance by forecast horizon, by month and horizon, and by location and horizon.

@@

The model is implemented in `stan` [@stan] and `R` (`4.2.0`) [@R] as an extension of the baseline model from the `forecast.vocs` R package (`0.0.9.7000`)  [@forecast.vocs].
We note that our use of a Heaviside step function introduces a discontinuity to the posterior making it less suited for use with `stan`. Other model formulations without this feature would be more efficient and robust.
The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
`cmstanr` surfaces several settings that trade-off between sampling speed and the robustness of the approach.
Here we take a conservative approach, as the model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum tree depth setting to 15. 
For real-time usage convergence was not assessed but during model development, the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergent transitions and exceedance of the maximum tree depth [@cmdstanr]. During development, posterior predictions were also visually compared to observed data.

@@

```{r vis-forecasts, fig.cap = "a.) Weighted interval scores at the 1-week and 4-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom. b.)Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one to four weeks). Incidence rates are shown on a log scale. 30\\%, 60\\%, and 90\\% credible intervals are shown. The black line and points are the notified cases as of the date of data extraction rather than those available at the time.", out.width = "95%", fig.height = 12, fig.width = 18}
p <- (
  plot_wis(wis, locs) +
   theme(strip.text.x = element_blank()) |
  plot_forecasts(forecasts, locs, ranges) +
    guides(fill = guide_none(), fill_ramp = guide_none(), col = guide_none())
) +
  plot_layout(widths = c(1, 4), guides = "collect") +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre"
  )
suppressWarnings(print(p))
```

@@

Evaluating the ensemble and surrogate models using the weighted interval score across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
Indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by approximately 100% on average with the surrogate model performing worst at the one and four week horizon.
However, much of this outperformance was driven by a small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a). 
If we instead consider median relative performance (note not a proper score) we find that, relative to the ensemble, the surrogate scored `r relative_median_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_median_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_median_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_median_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
Looking across the distribution of relative scores by horizon we see little evidence that relative forecast quality decreased with time, though the tail of the distribution appears slightly heavier at the 4 week time horizon (Figure \@ref(fig:assemble-eval) a).
`r relative_wis_thresholds[, better_than]` of surrogate forecasts scored better than the comparable ensemble forecast, `r relative_wis_thresholds[, fifty_percent]` performed within 50% of the comparable ensemble forecat, and scored `r relative_wis_thresholds[, more_than_100_percent]` performed worse than 100% more than the comparable ensemble forecast.

The surrogate models relative performance varied over time with sustantially worse performance in January to March compared to later in the year across all forecast horizons (Figure \@ref(fig:assemble-eval) b).
The majority in the difference in peformance appeared to be driven by a thicker right tail with this being particular feature of forecasts at longer horizons. Notably forecast performance in March had a bimodal distribution at the four week horizon with a substational fraction of surrogate forecasts outperforming the ensemble.
This variation in performance may have been linked to the BA.2 wave which peaked in most locations during this time period.

There was also substantial variation across forecast locations with the surrogate performing well in some locations at some forecast horizons, for example the four week horizon in the United Kingdom, and badly in others, for example the four week forecast in Switzerland (Figure \@ref(fig:assemble-eval) c).
In general, across locations, as observed overall, relative forecast performance was comparable across horizons though with a heavier right tail at longer horizons. Some locations showed less of this behaviour, for example Spain, and in some it was very dominant, for example Switzerland.

@@

```{r assemble-eval, fig.height = 12, fig.width = 12, out.width = "95%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale. a.) Density of the relative score by horizon. Horizontal black lines give the 5\\%, 35\\%, 65\\%, and 95\\% quantiles. b.) Density of the relative score by month for a given forecast horizon stratified by the one and four week forecast horizon. c.) Density of the relative score by forecast location stratified by the one and four week forecast horizon. The dashed line on all plots indicates when the ensemble forecast is equivalent to the surrogate forecast. The vertical black lines on the y axis give gives individual relative scores."}
p <- (
  (
    (
      plot_rwis_horizon / plot_rwis_by_month
    )  +
    plot_layout(heights = c(4, 5), guides = "collect")
  ) |
  plot_rwis_location
) +
  plot_layout(guides = "collect", widths = c(1, 1)) +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre"
  )
suppressWarnings(print(p))
```

@@

Overall the surrogate model was well calibrated at the 30%, 60% and 90% credible intervals whilst the ensemble model was in general poorly calibrated with calibration becoming worse at larger credible intervals.
When stratified by forecast horizon the ensemble forecast was best calibrated at the one week forecast horizon and then became progressively less well calibrated as foreast horizon increased with the worst calibration at the 90% credible interval (Figure \@ref(fig:plot-coverage) a). 
In comparison the surrogate forecast was less well calibrated at the one week forecast horizon with a tendancy to have a larger empirical coverage than required at longer horizons the surrogate forecast transioned to being well calibrated and then had a lower coverage than optimal though to a much lower degree than the ensemble forecast.
This was not the case for the 90% credible interval where the surrogate model stably covered slightly more than the expected interval indicating forecasts were somewhat overly uncertain for this intervals.

Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly precise at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
In contrast the surrogate ensemble forecast was generally equally well calibrated at shorter horizons with a tendency to produce conservative forecasts.
At longer horizons however quantiles below the median were liable to have poor coverage suggesting a bias towards over-predicting.

Breaking down the relative weighted interval score by forecast interval we observe that the majority of the difference in overall scores is likely driven by larger intervals and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) c).
This is true across forecast horizons but the magnitude of the different increases. This implies that the surogate and ensemble model produced comparably performing point forecasts and indeed if we consider the absolute error of the median forecasts we see that the median for the ensemble forecast was `r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` whilst the median for the surrogate forecast was `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`.
However, when we consider the mean absolute error of the median we find that the ensemble forecast outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs. `r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`) as noted elsewhere.
This indicates a right skew in performance for the surrogate model point forecasts not observed for the ensemble forecast which matches that observed when evaluating the full predictive distributions of both forecasts.

```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%", fig.cap = "a.) Empirical coverage of the surrogate and ensemble forecasts at the 90\\%, 60\\%, and 30\\% credible intervals stratified by forecast horizon. Ideally a well calibrated forecast should have empirical coverage for a given credible interval that equals the range of that interval (i.e 90\\%). b.) Median relative weighted interval score by quantile and forecast horizon for the surrogate forecast compared to the ensemble forecast. c.) Empirical coverage by quantile for both the surrrogate and ensemble forecasts. A well calibrated forecast should have empirical quantiles that match the theoretical ones. The green area of this figure corresponds to conservative forecasts."}
p <- (
  (
    (
      plot_coverage_range(coverage, ranges) /
      plot_rel_score_by_quantile(relative_interval_score)
    )  +
    plot_layout(heights = c(3, 1), guides = "collect")
  ) |
  plot_coverage_quantiles(scores)
) +
  plot_layout(widths = c(1, 2), guides = "collect") +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre"
  )
suppressWarnings(print(p))
```

@@

In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the ECDC forecast hub multi-team ensemble for forecasting test positive reported COVID-19 cases in ECDC nations and the UK.
We first defined a set of assumptions for how the surrogate model should behaved based on our observations of the ECDC forecast hub ensemble, our experience submitting forecasts to various forecast Hubs, and our observations of the performance of other models.
We restricted ourselves to a model that could be easily understood, that produced epidemiologically meaningful summary statistics, and that could be run with low compute resources.
We further provide a fully reproducibile workflow for running and evaluating this model using GitHub actions facilitating others to do the same.

Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble though with greater uncertainty.
In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased towards underprediction, and the surrrogate model appeared biased towards overprediction.
Evaluating the relative performance of the surrogate model compard to ECDC forecast hub ensemble we found that the mean performance was substantially worse but that this was largely driven by a relatively small subset of forecasts.
The median relative performance was only marginally worse across forecast horizons with the distribution of relative scores having a heavy right tail.
Performance varied by location and forecast date with the surrogate mdoel performing worse in the first part of 2021 which may have been linked to increasing indicence rates accross forecast locations linked to BA.2.
Unlike our earlier forecasting efforts the effect of outlier forecasts was much reduced though the surrogate model was less likely to outperform the ensemble than our earlier forecasting model methodology [@refNikosinGermany].
In general, the relative performance of the surrogate model was stable across forecast horizons, unlike our previous methods which rapidly became less performant, though the distribution of relative performance did have an increasingly heavy right tail as forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison the the hub ensemble.
As noted elsewhere, the forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger credible intervals whereas the surrogate model was substantially better calibrated for all targets though it tended to over cover at large intervals.
Stratifying the relative weighted interval score by interval we found that larger intervals (i.e the tails of the forecast distribution) were responsible for the majority of difference in performance difference. 
Indeed, median point performance of the forecasts were comparable though the surrogate model performed susbtantially worse on average due to its skewed performance distribution.

@@

We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders requirements. We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements. Our work maybe be useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to methodology developers thinking about how to develop forecasts that perform similary to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.

@@

All data and code is available here:

>>>>

Real-time evaluation has shown that each of these classes of models may perform well though in general semi-mechanistic models, and models that include human insights have outperformed other approaches during certain periods [@REF].

@@

The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@REF].

@@

Based on these insights from observing the performance of our forecast submissions and from the hub ensembles over time here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@REF] with aim of producing ensemble-like performance without sacrificing interpretability.
The first simplification we make is to model only weekly data, rather than using daily data and then aggregating, this means that the impact of daily reporting artefacts is mitigated.
It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model this leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.
The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
This reduces computational requirements and removes the need for external information on the delay from infection to report which can potentially be misspecified leading to potentially reduced forecast performance.
However, this sacrifices some of the interpretability of the forecast model as any epidemiological summary statistics we now calculate will be based on reported cases and not latent infections.
As discussed in [@gostic] this leads to varying amounts of bias depending on the epidemic phase. 
The final simplification we make is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@EpiNow2].
This represents a parsimonious approach in that we encode our expectation that the growth rate should vary over time and allow this to influence the forecast but we include only a single lag term, reducing the computational overhead of the model.
To model potential unobserved interventions and more general changes in transmission, we include an additional growth rate modifier restricted to be between 0 and 1 that differs depending on if the growth rate is positive or negative (due to potential differing responses when cases are growing or increasing) and that acts in a multiplicative fashion (meaning that larger absolute growth rates are reduced to zero growth more rapidly).
This reflects a simplified interpretation of how the ensemble appears to react to potential future changes by assuming a gradual return to stable incidence. 
The only observation for which we do not make an adaption is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion. 
This choice is motivated by our belief that the underlying transmission process is an exponential one and therefore a count error model (where variance is linked to the mean)  is a sensible choice though this is an area of current research.
We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations.

@@

  The ECDC forecast hub requests forecasts for one to four-week forecast horizon and requires forecasts to use a pre-specified format with 23 quantiles of the predictive probability distribution.
  No restrictions were placed on who could submit forecasts and the hub team actively invited participation from research groups known to be involved with COVID-19 forecasting projects.
  Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent-based models, semi-mechanistic models, expert elicitation models, and ensembles of multiple approaches [@https://covid19forecasthub.eu/community.html].
  Teams submitted forecasts at the latest two days after the complete dataset for the forecast week became available and were allowed to use all data available at the time of submission (i.e including up to two days of data for the current week). 
  The ensemble forecast was constructed by taking the median of all predictive quantiles without the exclusion of any validly submitted forecast (where validity was defined as passing minimal formatting checks by the hub team). 
An ensemble was only produced for locations with at least 3 independent forecast models including the hub baseline model.
  Submitted forecasts and target observations are available from the ECDC forecast hub GitHub repository [@REF]. We provide code in the repository of this study to streamline access.

@@

# Models coverage of forecast dates
forecast_dates_by_models <- metadata |>
  summarise_forecasts_by(var = "target_end_date", by = "model") |>
  DT(order(n)) |>
  DT(, n_per := n / forecast_dates) |>
  DT(, type := fcase(
      model %in% global_models, "global",
      model %in% local_models, "local",
      model %in% partial_models[, model], "partial",
      default = "other"
    )
  )
```

@@

We model the expectation ($\lambda_t$) of reported cases ($C_t$) as an order 1 autoregressive (AR(P)) process by epidemiological week ($t$).
The model is initialised by assuming that the initially reported cases are representative with a small amount of error (2.5%).
We assume a negative binomial observation model with overdispersion $\phi$ for reported cases ($C_t$).

@@

To standardise forecasts across forecast locations we first normalised both weekly notified test positive cases and forecast test positive cases by the population in the forecast region scaled to be an incidence rate per 10,000 people.
We then evaluated a subset of forecasts from the following countries Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom visually by forecast horizon (1-4 weeks) for both the ensemble and surrogate model. 
These countries were selected to include forecasts based on different numbers and types of submitted forecast models, to be at least partially representative of the full sample of forecast locations, and to include nations for which the authors had a good understanding of local data and transmission dynamics in the study period.

We evaluate forecasts for all locations and horizons quantitatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS).
Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the ECDC forecast hub  [@REF].
We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks).

To understand the relative performance of the surrogate model compared to the ensemble model, we calculate the relative performance (rWIS) by dividing the WIS for the surrogate model by the WIS of the ensemble model for all locations and forecast horizons.
To maintain the propriety of this score, we do this after first taking the means of scores for the relevant stratification.
We explore relative performance by forecast horizon, by month and horizon, and by location and horizon.

@@

The model is implemented in `stan` [@stan] and `R` (`4.2.0`) [@R] as an extension of the baseline model from the `forecast.vocs` R package (`0.0.9.7000`)  [@forecast.vocs].
We note that our use of a Heaviside step function introduces a discontinuity to the posterior making it less suited for use with `stan`. Other model formulations without this feature would be more efficient and robust.
The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
`cmstanr` surfaces several settings that trade off between sampling speed and the robustness of the approach.
Here we take a conservative approach, as the model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum tree depth setting to 15. 
For real-time usage convergence was not assessed but during model development, the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergent transitions and exceedance of the maximum tree depth [@cmdstanr]. During development, posterior predictions were also visually compared to observed data.

@@

```{r vis-forecasts, fig.cap = "a.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom. b.)Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one to four weeks). Incidence rates are shown on a log scale. 30\\%, 60\\%, and 90\\% credible intervals are shown. The black line and points are the notified cases as of the date of data extraction rather than those available at the time.", out.width = "95%", fig.height = 12, fig.width = 18}
p <- (
  plot_wis(wis, locs) +
   theme(strip.text.x = element_blank()) |
  plot_forecasts(forecasts, locs, ranges) +
    guides(fill = guide_none(), fill_ramp = guide_none(), col = guide_none())
) +
  plot_layout(widths = c(1, 4), guides = "collect") +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre"
  )
suppressWarnings(print(p))
```

@@

Evaluating the ensemble and surrogate models using the weighted interval score across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
Indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by approximately 100% on average with the surrogate model performing worst at the one and four-week horizon.
However, much of this outperformance was driven by a small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a). 
If we instead consider median relative performance (note not a proper score) we find that, relative to the ensemble, the surrogate scored `r relative_median_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_median_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_median_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_median_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
Looking across the distribution of relative scores by horizon we see little evidence that relative forecast quality decreased with time, though the tail of the distribution appears slightly heavier at the four-week time horizon (Figure \@ref(fig:assemble-eval) a).
`r relative_wis_thresholds[, better_than]` of surrogate forecasts scored better than the comparable ensemble forecast, `r relative_wis_thresholds[, fifty_percent]` performed within 50% of the comparable ensemble forecat, and scored `r relative_wis_thresholds[, more_than_100_percent]` performed worse than 100% more than the comparable ensemble forecast.

The surrogate model's relative performance varied over time with substantially worse performance from January to March compared to later in the year across all forecast horizons (Figure \@ref(fig:assemble-eval) b).
The majority of the difference in performance appeared to be driven by a thicker right tail with this being a particular feature of forecasts at longer horizons. Notably, forecast performance in March had a bimodal distribution at the four-week horizon with a substantial fraction of surrogate forecasts outperforming the ensemble.
This variation in performance may have been linked to the BA.2 wave which peaked in most locations during this period.

There was also substantial variation across forecast locations with the surrogate performing well in some locations at some forecast horizons, for example, the four-week horizon in the United Kingdom, and badly in others, for example, the four-week forecast in Switzerland (Figure \@ref(fig:assemble-eval) c).
In general, across locations, as observed overall, relative forecast performance was comparable across horizons though with a heavier right tail at longer horizons. Some locations showed less of this behaviour, for example, Spain, and in some, it was very dominant, for example, Switzerland.

@@

```{r assemble-eval, fig.height = 12, fig.width = 12, out.width = "95%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale. a.) The density of the relative score by horizon. Horizontal black lines give the 5\\%, 35\\%, 65\\%, and 95\\% quantiles. b.) The density of the relative score by month for a given forecast horizon stratified by the one and four-week forecast horizon. c.) The density of the relative score by forecast location stratified by the one and four-week forecast horizon. The dashed line on all plots indicates when the ensemble forecast is equivalent to the surrogate forecast. The vertical black lines on the y-axis give individual relative scores."}
p <- (
  (
    (
      plot_rwis_horizon / plot_rwis_by_month
    )  +
    plot_layout(heights = c(4, 5), guides = "collect")
  ) |
  plot_rwis_location
) +
  plot_layout(guides = "collect", widths = c(1, 1)) +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre"
  )
suppressWarnings(print(p))
```

@@

Overall the surrogate model was well calibrated at the 30%, 60% and 90% credible intervals whilst the ensemble model was in general poorly calibrated with calibration becoming worse at larger credible intervals.
When stratified by forecast horizon the ensemble forecast was best calibrated at the one-week forecast horizon and then became progressively less well calibrated as the forecast horizon increased with the worst calibration at the 90% credible interval (Figure \@ref(fig:plot-coverage) a). 
In comparison, the surrogate forecast was less well calibrated at the one-week forecast horizon with a tendency to have a larger empirical coverage than required at longer horizons the surrogate forecast transitioned to being well calibrated and then had lower coverage than optimal though to a much lower degree than the ensemble forecast.
This was not the case for the 90% credible interval where the surrogate model stably covered slightly more than the expected interval indicating forecasts were somewhat overly uncertain for this interval.

Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly precise at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
In contrast, the surrogate ensemble forecast was generally equally well calibrated at shorter horizons with a tendency to produce conservative forecasts.
At longer horizons, however, quantiles below the median were liable to have poor coverage suggesting a bias towards over-predicting.

Breaking down the relative weighted interval score by forecast interval we observe that the majority of the difference in overall scores is likely driven by larger intervals and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) c).
This is true across forecast horizons but the magnitude of the different increases. This implies that the surrogate and ensemble model produced comparably performing point forecasts and indeed if we consider the absolute error of the median forecasts we see that the median for the ensemble forecast was `r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` whilst the median for the surrogate forecast was `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`.
However, when we consider the mean absolute error of the median we find that the ensemble forecast outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs. `r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`) as noted elsewhere.
This indicates a right skew in performance for the surrogate model point forecasts not observed for the ensemble forecast which matches that observed when evaluating the full predictive distributions of both forecasts.

```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%", fig.cap = "a.) Empirical coverage of the surrogate and ensemble forecasts at the 90\\%, 60\\%, and 30\\% credible intervals stratified by forecast horizon. Ideally, a well-calibrated forecast should have empirical coverage for a given credible interval that equals the range of that interval (i.e 90\\%). b.) Median relative weighted interval score by quantile and forecast horizon for the surrogate forecast compared to the ensemble forecast. c.) Empirical coverage by quantile for both the surrogate and ensemble forecasts. A well-calibrated forecast should have empirical quantiles that match the theoretical ones. The green area of this figure corresponds to conservative forecasts."}
p <- (
  (
    (
      plot_coverage_range(coverage, ranges) /
      plot_rel_score_by_quantile(relative_interval_score)
    )  +
    plot_layout(heights = c(3, 1), guides = "collect")
  ) |
  plot_coverage_quantiles(scores)
) +
  plot_layout(widths = c(1, 2), guides = "collect") +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre"
  )
suppressWarnings(print(p))
```

@@

In this study, we defined a surrogate model aiming to replicate some of the observed behaviour of the ECDC forecast hub multi-team ensemble for forecasting test-positive reported COVID-19 cases in ECDC nations and the UK.
We first defined a set of assumptions for how the surrogate model should behave based on our observations of the ECDC forecast hub ensemble, our experience submitting forecasts to various forecast Hubs, and our observations of the performance of other models.
We restricted ourselves to a model that could be easily understood, that produced epidemiologically meaningful summary statistics, and that could be run with low compute resources.
We further provide a fully reproducible workflow for running and evaluating this model using GitHub actions facilitating others to do the same.

Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble though with greater uncertainty.
In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared biased towards overprediction.
Evaluating the relative performance of the surrogate model compared to the ECDC forecast hub ensemble we found that the mean performance was substantially worse but that this was largely driven by a relatively small subset of forecasts.
The median relative performance was only marginally worse across forecast horizons with the distribution of relative scores having a heavy right tail.
Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2021 which may have been linked to increasing incidence rates across forecast locations linked to BA.2.
Unlike our earlier forecasting efforts, the effect of outlier forecasts was much reduced though the surrogate model was less likely to outperform the ensemble than our earlier forecasting model methodology [@refNikosinGermany].
In general, the relative performance of the surrogate model was stable across forecast horizons, unlike our previous methods which rapidly became less performant, though the distribution of relative performance did have an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison the hub ensemble.
As noted elsewhere, the forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger credible intervals whereas the surrogate model was substantially better calibrated for all targets though it tended to overcover at large intervals.
Stratifying the relative weighted interval score by interval we found that larger intervals (i.e the tails of the forecast distribution) were responsible for the majority of the difference in performance. 
Indeed, the median point performance of the forecasts was comparable though the surrogate model performed substantially worse on average due to its skewed performance distribution.

@@

We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements. We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements. Our work may be useful for forecast users to understand the inherent assumptions of the forecasts they are making use of and to methodology developers thinking about how to develop forecasts that perform similary to current multi-model and multi-team forecast ensembles that are trusted by stakeholders.

@@

All data and code are available here:
